---
title: "IvsM_DESeq_multimapped_reads"
author: "Joanna Yeung"
date: "2025-08-12"
output: html_document
---

```{r}
hm_cols <- list(Condition=c("Interphase"="#96CDCD", "Mitotic"="#FF82AB"),
                Biorep=c("1"="#4269D0FF", "2"="#EFB118FF", "3"="#FF725CFF"),
                Centromere=c("Yes"="gold", "No"="white"))
# make TxDb object from Xenopus Laevis gtf file
TxDb.Xeno <- txdbmaker::makeTxDbFromGFF("/lustre/fs4/risc_lab/scratch/jyeung/ncbi-genomes-2022-09-11/Xenopus_laevis_v10.1/GCF_017654675.1_Xenopus_laevis_v10.1.anno.sorted.gff")

gene_annoXeno <- import("/lustre/fs4/risc_lab/scratch/jyeung/ncbi-genomes-2022-09-11/Xenopus_laevis_v10.1/GCF_017654675.1_Xenopus_laevis_v10.1.anno.sorted.gff")
```

```{r, warning=F}
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(chromVAR)
library(DESeq2)
library(pheatmap)
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(scales)
library(ChIPseeker)
library(pheatmap)
library(readr)
library(ggrepel)
library(chromPlot)
library(rtracklayer)
library(tidyr)
library(dplyr)
library(ggsci)
```

# import epigenetic info
```{r}
# import Centromeric regions
CENPA_OwenSmith_regions <- import.bed("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/ATAC_deep_sequencing_Novogene_10112024/peaks/GSE153058_xla_v10.2_cen.bed")

# import histone mods peaks
H3K9me3_peaks <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Sept2024/Xenbase/ChIPseq/peaks/H3K9me3_spermatozoon_adult_GSM1944484_1349_broadPeaks.bed")
H3K9me3_peaks <- GRanges(seqnames=H3K9me3_peaks$V1, ranges = IRanges(start=H3K9me3_peaks$V2, end=H3K9me3_peaks$V3), ID=H3K9me3_peaks$V4)

H3K4me3_peaks <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Sept2024/Xenbase/ChIPseq/peaks/H3K4me3_sperm_egg_extract_adult_frog_GSM3587294_15598_broadPeaks.bed")
H3K4me3_peaks <- GRanges(seqnames=H3K4me3_peaks$V1, ranges = IRanges(start=H3K4me3_peaks$V2, end=H3K4me3_peaks$V3), ID=H3K4me3_peaks$V4)

H3K27me3_peaks <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Sept2024/Xenbase/ChIPseq/peaks/H3K27me3_sperm_egg_extract_adult_frog_GSM3587306_15593_broadPeaks.bed")
H3K27me3_peaks <- GRanges(seqnames=H3K27me3_peaks$V1, ranges = IRanges(start=H3K27me3_peaks$V2, end=H3K27me3_peaks$V3), ID=H3K27me3_peaks$V4)
```

# Fragment Length distribution all reads
```{r}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_5_bam_FLD")
samplenames <- read_lines("samplenames.txt")
samplenames <- samplenames[c(1,3,4,2,5,6)]
metaData <- data.frame(sample=samplenames, Condition=c("Interphase", "Interphase", "Interphase", "Mitosis",  "Mitosis",  "Mitosis"), Genotype=rep("WT", length(samplenames)), Biorep=c("1", "2", "3", "1", "2", "3"))

samplesList <- list()
for(i in 1:length(samplenames)){
  samplesList[[i]] <- read.table(paste0(samplenames[i],"_allreads_insertdata.txt"), skip=10, header=T) # read in hist_data_withoutdups.log file for each sample
}
names(samplesList) <- samplenames

# want each sample to have the same value of insert sizes from 1-1000bp (we don't want missing insert sizes because then it is hard to plot on ggplot if different samples have different axes. We will fill in the insert sizes with no reads as NA. 
for(i in 1:length(samplesList)){
  hist <- data.frame(insert_size=1:1000) 
  samplesList[[i]] <- merge(hist, samplesList[[i]], by="insert_size", all=T)
  samplesList[[i]][is.na(samplesList[[i]]$All_Reads.fr_count), ]$All_Reads.fr_count <- 0  # convert NA values to 0
} 

final_hist <- do.call(rbind, samplesList) # rbind each samples' histogram dataframe into 1 final dataframe. 
final_hist$sample <- unlist(lapply(names(samplesList), rep, 1000)) # specify samples column so that we can color ggplot by sample.
final_hist$sample <- factor(final_hist$sample, levels=names(samplesList))
final_hist <- merge(final_hist, metaData, by="sample")

percent_hist <- samplesList

percent_count <-lapply(percent_hist, function(percent_hist){ percent_hist$All_Reads.fr_count/sum(na.omit(percent_hist$All_Reads.fr_count))})

for(i in 1:length(percent_hist)){
  percent_hist[[i]]$percent_count <- percent_count[[i]]
}
percent_hist <- do.call(rbind, percent_hist) # rbind each samples' histogram dataframe into 1 percent dataframe. 
percent_hist$sample <- unlist(lapply(names(samplesList), rep, 1000)) # specify samples column so that we can color ggplot by sample.
percent_hist$sample <- factor(percent_hist$sample, levels=names(samplesList))
percent_hist <- merge(percent_hist, metaData, by="sample")
```

# plot fragment length distribution
```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Manuscript_Figures/Figures"
pdf(paste0(pdf_dir, "/FLD.pdf"),  width=5, height=4)
ggplot(percent_hist, aes(x=insert_size, y=log(percent_count), colour=sample)) + 
  geom_line(linewidth=0.5) + 
  scale_colour_manual(values=c("#bfdede","#96CDCD","#528181", "#facbd9", "#FF82AB", "#a6496a")) +
   theme_bw(base_size = 10) + 
    scale_x_continuous(n.breaks=10)+
  ylab("log10(Percent total fragments)") + 
  xlab("Fragment length (bp)")+
  theme_classic()+
  theme(legend.position="bottom")

ggplot(percent_hist, aes(x=insert_size, y=percent_count, colour=sample)) + 
  geom_line(linewidth=0.5) + 
  scale_colour_manual(values=c("#bfdede","#96CDCD","#528181", "#facbd9", "#FF82AB", "#a6496a")) +
   theme_bw(base_size = 10) + 
   scale_x_continuous(n.breaks=10)+
  ylab("Percent total fragments") + 
    xlab("Fragment length (bp)")+
    theme_classic()+
  theme(legend.position="bottom")
dev.off()
```


# featureCounts, PCA & Correlation including HIRA IvsM counts under IvsM masterpeaks. 
```{r}
setwd("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_7_featureCounts")
IvsM_samplenames <- dir(pattern="atacHAK")[c(1,3,4,2,5,6)]

# make metaData from all samples
IvsM.metaData <- data.frame(sample=IvsM_samplenames, Condition=c("Interphase", "Interphase", "Interphase", "Mitotic",  "Mitotic",  "Mitotic"), Genotype=c(rep("Interphase", 3), rep("Mitosis", 3)), Biorep=c("1", "2", "3", "1", "2", "3"))

feature_IvsMCounts <- list()
for(i in 1:length(IvsM_samplenames)){
  feature_IvsMCounts[[i]] <- read.table(paste0(IvsM_samplenames[i],"/", IvsM_samplenames[i], "_allOverlaps_counts.txt"), skip=1, header=T) # read in feature_IvsMCounts file for each sample
  colnames(feature_IvsMCounts[[i]])[ncol(feature_IvsMCounts[[i]])] <- IvsM_samplenames[i]
}
names(feature_IvsMCounts) <- IvsM_samplenames

IvsMCounts <- matrix(nrow=nrow(feature_IvsMCounts[[1]]), ncol=length(feature_IvsMCounts), dimnames=list(feature_IvsMCounts[[1]]$Geneid, names(feature_IvsMCounts)))
for(i in 1:length(feature_IvsMCounts)){
  IvsMCounts[ ,i] <- as.integer(feature_IvsMCounts[[i]][, 7])
}

# make DESeq object from frag_IvsMCounts matrix
IvsM.dds <- DESeq2::DESeqDataSetFromMatrix(countData=IvsMCounts, design=~Biorep+Condition, colData= IvsM.metaData)
IvsM.dds <- estimateSizeFactors(IvsM.dds)
IvsM.dds <- estimateDispersions(IvsM.dds)


# rlog normalize for QC metrics visualization
rlog_IvsM.dds <- rlog(IvsM.dds)
```

# plot sample correlation matrix
```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Manuscript_Figures/Figures"

# get pearson correlation coefficients for pairwise comparisons between samples
sampleDists <- dist(t(assay(rlog_IvsM.dds ))) # get rlog normalized count matrix & convert to distance measure for calculating correlations. 
sampleDistMatrix <- as.matrix(sampleDists)
row.names(sampleDistMatrix) <- IvsM.metaData$sample
# plot sample distance matrix

row.names(IvsM.metaData) <- IvsM.metaData$sample
pdf(paste0(pdf_dir, "/sampleDistMatrix.pdf"), width=5, height=4)
pheatmap(sampleDistMatrix, colorRampPalette(rev(c("white","white", "#FFFFB2","#FED976", "#FEB24C", "#FD8D3C", "#E31A1C", "#B10026")))(100), annotation_row = IvsM.metaData[ ,-1], annotation_col = IvsM.metaData[ ,-1], annotation_colors = hm_cols) # plot correlation heatmap
dev.off()
```

# plot PCA:
```{r}
# PCA based on all peaks
scaled_data <- scale(t(counts(dds, normalized=T)))
pca_result <- prcomp(scaled_data)

# plot the variance explained by each PC
plot(1:length(pca_result$sdev), pca_result$sdev^2 / sum(pca_result$sdev^2),
     type = "b", xlab = "Principal Component", ylab = "Variance Explained")

# pca_result$x is where the sample is located in PCA space. 
pca_result_df <- as.data.frame(pca_result$x)

pca_result_df <- cbind(pca_result_df, IvsM.metaData) # include variables associated with samples

VarianceExplained <- as.integer((pca_result$sdev^2 / sum(pca_result$sdev^2))*100) # variance explained by each PC 
# PC1 vs PC2

pdf(paste0(pdf_dir, "/PCA.pdf"), width=5, height=3)
ggplot(data=pca_result_df) + geom_point(aes(x=PC1, y=PC2, color=Condition, shape=Biorep), size=3)+geom_text_repel(aes(x=PC1, y=PC2, label=sample))+xlab(paste0("PC1: ", VarianceExplained[1], "% Variance Explained"))+ylab(paste0("PC2: ", VarianceExplained[2], "% Variance Explained"))+theme_classic()+
  ggtitle("PCA: Normalized counts, all Peaks")+
  scale_colour_manual(values=hm_cols$Condition)+
  scale_x_continuous(limits = c(-700,500))+
  scale_y_continuous(limits = c(-700,500))

# by treatment- top 500 most variable peaks
pca_results_top500 <- plotPCA(rlog_IvsM.dds, intgroup=c("Condition", "Biorep"), returnData=TRUE)

plotPCA(rlog_IvsM.dds, intgroup=c("Condition", "Biorep"))
ggplot(data=pca_results_top500) + geom_point(aes(x=PC1, y=PC2, color=Condition, shape=Biorep), size=3)+geom_text_repel(aes(x=PC1, y=PC2, label=name))+
  xlab(paste0("PC1: 96% Variance Explained"))+
  ylab(paste0("PC2: 2% Variance Explained"))+
  theme_classic()+
  ggtitle("PCA: Rlog normalized counts, top 500 most variable peaks")+
  scale_colour_manual(values=hm_cols$Condition)+
  scale_x_continuous(limits = c(-30,30))+
  scale_y_continuous(limits = c(-30,30))
dev.off()
```

# perform DESeq-- significant results are absolute FC > 2.46 (log2FC > 1.3) & padj < 0.05
```{r}
IvsM.dds <- DESeq(IvsM.dds)
IvsM.res <- results(IvsM.dds)
#write.csv(IvsM.res, "/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/post_peakcalling_analysis/Figures/IvsM.DESeq.res.csv")
# get significant results
IvsM.sigResWald <- IvsM.res[IvsM.res$padj < 0.05 & abs(IvsM.res$log2FoldChange) > 1.3, ]
IvsM.sigResWald <- IvsM.sigResWald[order(IvsM.sigResWald$log2FoldChange, decreasing=T), ]
# get peak_ids of significant peaks. 
sigPeak.ID.Wald <- unique(row.names(IvsM.sigResWald))
```

# rlogCounts under significant peaks
```{r}
rlog_IvsM.Mat <- assays(IvsM.dds)[[1]]
rlog_IvsM.sigMat <- rlog_IvsM.Mat[row.names(rlog_IvsM.Mat) %in% sigPeak.ID.Wald, ]

set.seed(2025)
k <- pheatmap(rlog_IvsM.sigMat, scale="row", k=2)

# annotate which cluster peaks belong to 
anno.IvsM.sigPeaks_df <- data.frame(row.names = names(k$kmeans$cluster), Cluster=k$kmeans$cluster)
# annotate if peaks overlap with centromeres
anno.IvsM.sigPeaks_df$Centromere <- ifelse(row.names(anno.IvsM.sigPeaks_df) %in% IvsM.masterpeaks[IvsM.masterpeaks %over% CENPA_OwenSmith_regions]$peak_id, "Yes", "No")

# Cluster 2= Interphase Up peaks
# Cluster 1= Mitosis Up peaks
```

# get all peaks
```{r}
# get GRanges of all peaks
IvsM.masterpeaks <- import("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_6_IDR_tagalign/nonredundant_masterpeaks_IDR.900000.0.1.gtf")

Interphase_peaks <- import("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_6_IDR_tagalign/Interphase/Interphase_REP1_VS_REP2_VS_REP3.IDR.900000.0.1.narrowPeak")
Interphase_peaks <- GenomicRanges::reduce(Interphase_peaks)
Interphase_peaks <- IvsM.masterpeaks[IvsM.masterpeaks %over% Interphase_peaks]

Mitosis_peaks <- import("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_6_IDR_tagalign/Mitosis/Mitosis_REP1_VS_REP2_VS_REP3.IDR.900000.0.1.narrowPeak")
Mitosis_peaks <- GenomicRanges::reduce(Mitosis_peaks)
Mitosis_peaks <- IvsM.masterpeaks[IvsM.masterpeaks %over% Mitosis_peaks]

# all ATAC-seq peaks that overlap with centromere regions defined by Smith et al, 2021. 
CENP_peaks <- IvsM.masterpeaks[IvsM.masterpeaks %over% CENPA_OwenSmith_regions]
export.bed(CENP_peaks[ ,-3], con="/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/ATAC_deep_sequencing_Novogene_10112024/peaks/IvsM.masterpeaks_over_CENPA_GSE153058_xla_v10.2_cen.bed")
```

# get significant peaks
```{r}
# get GRanges of sigPeaks
IvsM.sigResWald_GR <- IvsM.masterpeaks[IvsM.masterpeaks$peak_id %in% sigPeak.ID.Wald]
IvsM.sigResWald_GR <- GRangesList(All=IvsM.sigResWald_GR,
InterphaseUp=IvsM.sigResWald_GR[IvsM.sigResWald_GR$peak_id %in% names(which(k$kmeans$cluster ==2))],
MitosisUp=IvsM.sigResWald_GR[IvsM.sigResWald_GR$peak_id %in% names(which(k$kmeans$cluster ==1))] )

anno.IvsM.sigResWald_GR <- lapply(IvsM.sigResWald_GR, annotatePeak, TxDb = TxDb.Xeno)
anno.IvsM.masterpeaks_GR <- annotatePeak(IvsM.masterpeaks, TxDb = TxDb.Xeno)
```

# plot Overlap of significant peaks with centromeric regions (Venn diagram)
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2),
      proportional = FALSE)
}
pdf(paste0(pdf_dir, "/IvsM.diffsigPeaks.centromere.VennDiagram.pdf"))
create_proportional_venn(sets=list(IvsM.sigResWald_GR$InterphaseUp$peak_id, 
                                   IvsM.sigResWald_GR$MitosisUp$peak_id,
                                   CENP_peaks$peak_id, 
                        Interphase_peaks$peak_id, 
                        Mitosis_peaks$peak_id ),
                         set_labels = c("Interphase Up", "Mitotic Up", "Centromeres", "all Interphase peaks", 
                                        "all Mitosis peaks"), 
                         colors = c("#96CDCD", "#FF82AB", "gold", "#d6edea", "#f7d5e1" )
)

create_proportional_venn(sets=list(IvsM.sigResWald_GR$InterphaseUp$peak_id, 
                                   IvsM.sigResWald_GR$MitosisUp$peak_id,
                                   CENP_peaks$peak_id, 
                        IvsM.masterpeaks$peak_id),
                         set_labels = c("Interphase Up", "Mitotic Up", "Centromeres",
                                        "All accessible peaks"), 
                         colors = c("#96CDCD", "#FF82AB", "gold", "lightgrey")
)
dev.off()

```

# merge bams from technical replicates from each condition, then convert into merged bigwigs
```{bash}
sbatch -p hpc_bigmem_b /lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Manuscript_Figures/merge_bams.slurm
```

```{r}
```


# extract annotation info of significant peaks
```{r}
anno.IvsM.sigResWald_GR <- lapply(anno.IvsM.sigResWald_GR, as.GRanges)

for(i in 1:length(anno.IvsM.sigResWald_GR)){
anno.IvsM.sigResWald_GR[[i]]$gene_biotype <- gene_annoXeno[match(anno.IvsM.sigResWald_GR[[i]]$geneId, gene_annoXeno$gene_id), ]$gene_biotype
anno.IvsM.sigResWald_GR[[i]]$gbkey <- gene_annoXeno[match(anno.IvsM.sigResWald_GR[[i]]$geneId, gene_annoXeno$gene_id), ]$gbkey
anno.IvsM.sigResWald_GR[[i]]$annotation <- gsub("^\\s*(\\S+)\\s(?!UTR).*", "\\1", anno.IvsM.sigResWald_GR[[i]]$annotation, perl = TRUE)
anno.IvsM.sigResWald_GR[[i]]$Centromere <- ifelse(anno.IvsM.sigResWald_GR[[i]] %over% CENPA_OwenSmith_regions, "Centromeric", "Non-centromeric")
}

# convert anno GR of sig peaks to df 
anno.IvsM.sigResWald_df <- lapply(anno.IvsM.sigResWald_GR, as.data.frame)
anno.IvsM.sigResWald_df$InterphaseUp$SigPeak.Cat <- "Interphase Up"
anno.IvsM.sigResWald_df$MitosisUp$SigPeak.Cat <- "Mitosis Up"
anno.IvsM.sigResWald_df <- rbind(anno.IvsM.sigResWald_df$InterphaseUp, anno.IvsM.sigResWald_df$MitosisUp)
```

# plot type of nearest gene for sig peaks (not much difference I vs M)
```{r}
IvsM.sig.GeneAnno_df <- anno.IvsM.sigResWald_df %>%
  group_by(SigPeak.Cat, gene_biotype, annotation) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(SigPeak.Cat) %>%
  mutate(PercentTotalPeaks = (n / sum(n))*100, 
         NumPeaks = n)
# PercentTotalPeaks=divide by percent total peaks within each SigPeak.Cat which would be either Interphase Up or Mitosis Up. 
#NumPeaks=total number of peaks within SigPeak.Cat, gene_biotype, & annotation group. 

pdf(paste0(pdf_dir, "/NearestGeneBiotype.IvsM.pdf"), height=4)
ggplot(IvsM.sig.GeneAnno_df, aes(x=gene_biotype, y=PercentTotalPeaks, fill=annotation))+
  geom_col()+
  facet_wrap(~SigPeak.Cat)+
theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("all types of nearest genes")+
   ylab("% Total Peaks within Condition")+
  scale_fill_cosmic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

# by percent total peaks within condition
ggplot(IvsM.sig.GeneAnno_df[!IvsM.sig.GeneAnno_df$gene_biotype %in% c("guide_RNA", "protein_coding", "pseudogene", "lncRNA"), ], aes(x=gene_biotype, y=PercentTotalPeaks, fill=annotation))+
  geom_col()+
  facet_wrap(~SigPeak.Cat)+
theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Type of nearest gene: least abundant")+
  ylab("% Total Peaks within Condition")+
  scale_fill_cosmic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

# by number of peaks within condition
ggplot(IvsM.sig.GeneAnno_df[!IvsM.sig.GeneAnno_df$gene_biotype %in% c("guide_RNA", "protein_coding", "pseudogene", "lncRNA"), ], aes(x=gene_biotype, y=NumPeaks, fill=annotation))+
  geom_col()+
  facet_wrap(~SigPeak.Cat)+
theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Type of nearest gene: least abundant")+
  ylab("Number of Peaks")+
  scale_fill_cosmic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

# by percent total peaks within condition
ggplot(IvsM.sig.GeneAnno_df[IvsM.sig.GeneAnno_df$gene_biotype %in% c("guide_RNA", "protein_coding", "pseudogene", "lncRNA"), ], aes(x=gene_biotype, y=PercentTotalPeaks, fill=annotation))+
  geom_col()+
  facet_wrap(~SigPeak.Cat)+
theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Type of nearest gene: most abundant")+
  ylab("% Total Peaks within Condition")+
  scale_fill_cosmic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

# by number of peaks within condition
ggplot(IvsM.sig.GeneAnno_df[IvsM.sig.GeneAnno_df$gene_biotype %in% c("guide_RNA", "protein_coding", "pseudogene", "lncRNA"), ], aes(x=gene_biotype, y=NumPeaks, fill=annotation))+
  geom_col()+
  facet_wrap(~SigPeak.Cat)+
theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Type of nearest gene: most abundant")+
  ylab("Number of Peaks")+
  scale_fill_cosmic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
dev.off()
```

# extract normalized counts under significant peaks
```{r}
IvsM.NormCountsMat <- counts(IvsM.dds, normalized=T)
sig.IvsM.NormCounts <- IvsM.NormCountsMat[row.names(IvsM.NormCountsMat) %in% sigPeak.ID.Wald, ]
# convert matrix of normalized counts of sig peaks into a dataframe with metaData information for ggplot to accept
sig.IvsM.NormCounts_df <- list()
for(i in 1:nrow(sig.IvsM.NormCounts)){
sig.IvsM.NormCounts_df[[i]] <- data.frame(Counts=sig.IvsM.NormCounts[i, ], sample=names((sig.IvsM.NormCounts[i, ])), Peak_ID= row.names(sig.IvsM.NormCounts)[i], IvsM.metaData)
}
sig.IvsM.NormCounts_df <- do.call(rbind, sig.IvsM.NormCounts_df) # combine list into 1 giant dataframe
```

# convert rlogCounts matrix of significant peaks to df. 
```{r}
sig.IvsM.rlogCounts_df <- list()
for(i in 1:nrow(rlog_IvsM.sigMat)){
sig.IvsM.rlogCounts_df[[i]] <- data.frame(Counts=rlog_IvsM.sigMat[i, ], sample=names((rlog_IvsM.sigMat[i, ])), Peak_ID= row.names(rlog_IvsM.sigMat)[i], IvsM.metaData)
}
sig.IvsM.rlogCounts_df <- do.call(rbind, sig.IvsM.rlogCounts_df) # combine list into 1 giant dataframe
```
