---
title: "IvsM_DESeq_multimapped_reads"
author: "Joanna Yeung"
date: "2025-08-12"
output: html_document
---

```{r}
hm_cols <- list(Condition=c("Interphase"="#90CAF8FF", "Mitotic"="#E7210A"),
                Biorep=c("1"="#4269D0FF", "2"="#EFB118FF", "3"="#FF725CFF"),
                Centromere=c("Yes"="gold", "No"="white"))
# make TxDb object from Xenopus Laevis gtf file
TxDb.Xeno <- txdbmaker::makeTxDbFromGFF("/lustre/fs4/risc_lab/scratch/jyeung/ncbi-genomes-2022-09-11/Xenopus_laevis_v10.1/GCF_017654675.1_Xenopus_laevis_v10.1.anno.sorted.gff")

gene_annoXeno <- import("/lustre/fs4/risc_lab/scratch/jyeung/ncbi-genomes-2022-09-11/Xenopus_laevis_v10.1/GCF_017654675.1_Xenopus_laevis_v10.1.anno.sorted.gff")
```

```{r, warning=F}
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(chromVAR)
library(DESeq2)
library(pheatmap)
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(scales)
library(ChIPseeker)
library(pheatmap)
library(readr)
library(ggrepel)
library(chromPlot)
library(rtracklayer)
library(tidyr)
library(dplyr)
library(ggsci)
```

# import epigenetic info
```{r}
# import Centromeric regions
CENPA_OwenSmith_regions <- import.bed("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/ATAC_deep_sequencing_Novogene_10112024/peaks/GSE153058_xla_v10.2_cen.bed")

# import histone mods peaks
H3K9me3_peaks <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Sept2024/Xenbase/ChIPseq/peaks/H3K9me3_spermatozoon_adult_GSM1944484_1349_broadPeaks.bed")
H3K9me3_peaks <- GRanges(seqnames=H3K9me3_peaks$V1, ranges = IRanges(start=H3K9me3_peaks$V2, end=H3K9me3_peaks$V3), ID=H3K9me3_peaks$V4)

H3K4me3_peaks <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Sept2024/Xenbase/ChIPseq/peaks/H3K4me3_sperm_egg_extract_adult_frog_GSM3587294_15598_broadPeaks.bed")
H3K4me3_peaks <- GRanges(seqnames=H3K4me3_peaks$V1, ranges = IRanges(start=H3K4me3_peaks$V2, end=H3K4me3_peaks$V3), ID=H3K4me3_peaks$V4)

H3K27me3_peaks <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Sept2024/Xenbase/ChIPseq/peaks/H3K27me3_sperm_egg_extract_adult_frog_GSM3587306_15593_broadPeaks.bed")
H3K27me3_peaks <- GRanges(seqnames=H3K27me3_peaks$V1, ranges = IRanges(start=H3K27me3_peaks$V2, end=H3K27me3_peaks$V3), ID=H3K27me3_peaks$V4)
```

# Fragment Length distribution all reads
```{r}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_5_bam_FLD")
samplenames <- read_lines("samplenames.txt")
samplenames <- samplenames[c(1,3,4,2,5,6)]
metaData <- data.frame(sample=samplenames, Condition=c("Interphase", "Interphase", "Interphase", "Mitosis",  "Mitosis",  "Mitosis"), Genotype=rep("WT", length(samplenames)), Biorep=c("1", "2", "3", "1", "2", "3"))

samplesList <- list()
for(i in 1:length(samplenames)){
  samplesList[[i]] <- read.table(paste0(samplenames[i],"_allreads_insertdata.txt"), skip=10, header=T) # read in hist_data_withoutdups.log file for each sample
}
names(samplesList) <- samplenames

# want each sample to have the same value of insert sizes from 1-1000bp (we don't want missing insert sizes because then it is hard to plot on ggplot if different samples have different axes. We will fill in the insert sizes with no reads as NA. 
for(i in 1:length(samplesList)){
  hist <- data.frame(insert_size=1:1000) 
  samplesList[[i]] <- merge(hist, samplesList[[i]], by="insert_size", all=T)
  samplesList[[i]][is.na(samplesList[[i]]$All_Reads.fr_count), ]$All_Reads.fr_count <- 0  # convert NA values to 0
} 

final_hist <- do.call(rbind, samplesList) # rbind each samples' histogram dataframe into 1 final dataframe. 
final_hist$sample <- unlist(lapply(names(samplesList), rep, 1000)) # specify samples column so that we can color ggplot by sample.
final_hist$sample <- factor(final_hist$sample, levels=names(samplesList))
final_hist <- merge(final_hist, metaData, by="sample")

percent_hist <- samplesList

percent_count <-lapply(percent_hist, function(percent_hist){ percent_hist$All_Reads.fr_count/sum(na.omit(percent_hist$All_Reads.fr_count))})

for(i in 1:length(percent_hist)){
  percent_hist[[i]]$percent_count <- percent_count[[i]]
}
percent_hist <- do.call(rbind, percent_hist) # rbind each samples' histogram dataframe into 1 percent dataframe. 
percent_hist$sample <- unlist(lapply(names(samplesList), rep, 1000)) # specify samples column so that we can color ggplot by sample.
percent_hist$sample <- factor(percent_hist$sample, levels=names(samplesList))
percent_hist <- merge(percent_hist, metaData, by="sample")
```

# plot fragment length distribution
```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Manuscript_Figures/Figures"
pdf(paste0(pdf_dir, "/FLD.pdf"),  width=5, height=4)
ggplot(percent_hist, aes(x=insert_size, y=log(percent_count), colour=sample)) + 
  geom_line(linewidth=0.5) + 
  scale_colour_manual(values=c("#BADEFAFF","#90CAF8FF","#64B4F6FF", "#FC9272", "#EF3B2C", "#A50F15")) +
   theme_bw(base_size = 10) + 
    scale_x_continuous(n.breaks=10)+
  ylab("log10(Percent total fragments)") + 
  xlab("Fragment length (bp)")+
  theme_classic()+
  theme(legend.position="bottom")

ggplot(percent_hist, aes(x=insert_size, y=percent_count, colour=sample)) + 
  geom_line(linewidth=0.5) + 
  scale_colour_manual(values=c("#BADEFAFF", "#90CAF8FF", "#64B4F6FF", "#FC9272", "#EF3B2C", "#A50F15")) +
   theme_bw(base_size = 10) + 
   scale_x_continuous(n.breaks=10)+
  ylab("Percent total fragments") + 
    xlab("Fragment length (bp)")+
    theme_classic()+
  theme(legend.position="bottom")
dev.off()
```


# featureCounts, PCA & Correlation including HIRA IvsM counts under IvsM masterpeaks. 
```{r}
setwd("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_7_featureCounts")
IvsM_samplenames <- dir(pattern="atacHAK")[c(1,3,4,2,5,6)]

# make metaData from all samples
IvsM.metaData <- data.frame(sample=IvsM_samplenames, Condition=c("Interphase", "Interphase", "Interphase", "Mitotic",  "Mitotic",  "Mitotic"), Genotype=c(rep("Interphase", 3), rep("Mitosis", 3)), Biorep=c("1", "2", "3", "1", "2", "3"))

feature_IvsMCounts <- list()
for(i in 1:length(IvsM_samplenames)){
  feature_IvsMCounts[[i]] <- read.table(paste0(IvsM_samplenames[i],"/", IvsM_samplenames[i], "_allOverlaps_counts.txt"), skip=1, header=T) # read in feature_IvsMCounts file for each sample
  colnames(feature_IvsMCounts[[i]])[ncol(feature_IvsMCounts[[i]])] <- IvsM_samplenames[i]
}
names(feature_IvsMCounts) <- IvsM_samplenames

IvsMCounts <- matrix(nrow=nrow(feature_IvsMCounts[[1]]), ncol=length(feature_IvsMCounts), dimnames=list(feature_IvsMCounts[[1]]$Geneid, names(feature_IvsMCounts)))
for(i in 1:length(feature_IvsMCounts)){
  IvsMCounts[ ,i] <- as.integer(feature_IvsMCounts[[i]][, 7])
}

# make DESeq object from frag_IvsMCounts matrix
IvsM.dds <- DESeq2::DESeqDataSetFromMatrix(countData=IvsMCounts, design=~Biorep+Condition, colData= IvsM.metaData)
IvsM.dds <- estimateSizeFactors(IvsM.dds)
IvsM.dds <- estimateDispersions(IvsM.dds)


# rlog normalize for QC metrics visualization
rlog_IvsM.dds <- rlog(IvsM.dds)
```

# plot sample correlation matrix
```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/for_Hide/Manuscript_Figures/Figures"

# get pearson correlation coefficients for pairwise comparisons between samples
sampleDists <- dist(t(assay(rlog_IvsM.dds ))) # get rlog normalized count matrix & convert to distance measure for calculating correlations. 
sampleDistMatrix <- as.matrix(sampleDists)
row.names(sampleDistMatrix) <- IvsM.metaData$sample
# plot sample distance matrix

row.names(IvsM.metaData) <- IvsM.metaData$sample
pdf(paste0(pdf_dir, "/sampleDistMatrix.pdf"), width=5, height=4)
pheatmap(sampleDistMatrix, colorRampPalette(rev(c("white","white", "#FFFFB2","#FED976", "#FEB24C", "#FD8D3C", "#E31A1C", "#B10026")))(100), annotation_row = IvsM.metaData[ ,-1], annotation_col = IvsM.metaData[ ,-1], annotation_colors = hm_cols) # plot correlation heatmap
dev.off()
```

# plot PCA:
```{r}
# PCA based on all peaks
scaled_data <- scale(t(counts(dds, normalized=T)))
pca_result <- prcomp(scaled_data)

# plot the variance explained by each PC
plot(1:length(pca_result$sdev), pca_result$sdev^2 / sum(pca_result$sdev^2),
     type = "b", xlab = "Principal Component", ylab = "Variance Explained")

# pca_result$x is where the sample is located in PCA space. 
pca_result_df <- as.data.frame(pca_result$x)

pca_result_df <- cbind(pca_result_df, IvsM.metaData) # include variables associated with samples

VarianceExplained <- as.integer((pca_result$sdev^2 / sum(pca_result$sdev^2))*100) # variance explained by each PC 
# PC1 vs PC2

pdf(paste0(pdf_dir, "/PCA.pdf"), width=5, height=3)
ggplot(data=pca_result_df) + geom_point(aes(x=PC1, y=PC2, color=Condition, shape=Biorep), size=3)+geom_text_repel(aes(x=PC1, y=PC2, label=sample))+xlab(paste0("PC1: ", VarianceExplained[1], "% Variance Explained"))+ylab(paste0("PC2: ", VarianceExplained[2], "% Variance Explained"))+theme_classic()+
  ggtitle("PCA: Normalized counts, all Peaks")+
  scale_colour_manual(values=hm_cols$Condition)+
  scale_x_continuous(limits = c(-700,500))+
  scale_y_continuous(limits = c(-700,500))

# by treatment- top 500 most variable peaks
pca_results_top500 <- plotPCA(rlog_IvsM.dds, intgroup=c("Condition", "Biorep"), returnData=TRUE)

plotPCA(rlog_IvsM.dds, intgroup=c("Condition", "Biorep"))
ggplot(data=pca_results_top500) + geom_point(aes(x=PC1, y=PC2, color=Condition, shape=Biorep), size=3)+geom_text_repel(aes(x=PC1, y=PC2, label=name))+
  xlab(paste0("PC1: 96% Variance Explained"))+
  ylab(paste0("PC2: 2% Variance Explained"))+
  theme_classic()+
  ggtitle("PCA: Rlog normalized counts, top 500 most variable peaks")+
  scale_colour_manual(values=hm_cols$Condition)+
  scale_x_continuous(limits = c(-30,30))+
  scale_y_continuous(limits = c(-30,30))
dev.off()
```

# perform DESeq-- significant results are absolute FC > 2.46 (log2FC > 1.3) & padj < 0.05
```{r}
IvsM.dds <- DESeq(IvsM.dds)
IvsM.res <- results(IvsM.dds)
#write.csv(IvsM.res, "/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/post_peakcalling_analysis/Figures/IvsM.DESeq.res.csv")
# get significant results
IvsM.sigResWald <- IvsM.res[IvsM.res$padj < 0.05 & abs(IvsM.res$log2FoldChange) > 1.3, ]
IvsM.sigResWald <- IvsM.sigResWald[order(IvsM.sigResWald$log2FoldChange, decreasing=T), ]
# get peak_ids of significant peaks. 
sigPeak.ID.Wald <- unique(row.names(IvsM.sigResWald))
```

# rlogCounts under significant peaks
```{r}
rlog_IvsM.Mat <- assays(IvsM.dds)[[1]]
rlog_IvsM.sigMat <- rlog_IvsM.Mat[row.names(rlog_IvsM.Mat) %in% sigPeak.ID.Wald, ]

set.seed(2025)
k <- pheatmap(rlog_IvsM.sigMat, scale="row", k=2)

# annotate which cluster peaks belong to 
anno.IvsM.sigPeaks_df <- data.frame(row.names = names(k$kmeans$cluster), Cluster=k$kmeans$cluster)
# annotate if peaks overlap with centromeres
anno.IvsM.sigPeaks_df$Centromere <- ifelse(row.names(anno.IvsM.sigPeaks_df) %in% IvsM.masterpeaks[IvsM.masterpeaks %over% CENPA_OwenSmith_regions]$peak_id, "Yes", "No")

# Cluster 2= Interphase Up peaks
# Cluster 1= Mitosis Up peaks
```

# get all peaks
```{r}
# get GRanges of all peaks
IvsM.masterpeaks <- import("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_6_IDR_tagalign/nonredundant_masterpeaks_IDR.900000.0.1.gtf")

Interphase_peaks <- import("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_6_IDR_tagalign/Interphase/Interphase_REP1_VS_REP2_VS_REP3.IDR.900000.0.1.narrowPeak")
Interphase_peaks <- GenomicRanges::reduce(Interphase_peaks)
Interphase_peaks <- IvsM.masterpeaks[IvsM.masterpeaks %over% Interphase_peaks]

Mitosis_peaks <- import("/lustre/fs4/fnbk_lab/scratch/hkonishi/spermATAC/ATAC_deep_sequencing_Novogene_10112024/Justin_ATAC_pipeline/atac_6_IDR_tagalign/Mitosis/Mitosis_REP1_VS_REP2_VS_REP3.IDR.900000.0.1.narrowPeak")
Mitosis_peaks <- GenomicRanges::reduce(Mitosis_peaks)
Mitosis_peaks <- IvsM.masterpeaks[IvsM.masterpeaks %over% Mitosis_peaks]
```

# get significant peaks
```{r}
# get GRanges of sigPeaks
IvsM.sigResWald_GR <- IvsM.masterpeaks[IvsM.masterpeaks$peak_id %in% sigPeak.ID.Wald]
IvsM.sigResWald_GR <- GRangesList(All=IvsM.sigResWald_GR,
InterphaseUp=IvsM.sigResWald_GR[IvsM.sigResWald_GR$peak_id %in% names(which(k$kmeans$cluster ==2))],
MitosisUp=IvsM.sigResWald_GR[IvsM.sigResWald_GR$peak_id %in% names(which(k$kmeans$cluster ==1))] )
```

# plot Overlap of significant peaks with centromeric regions (Venn diagram)
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2),
      proportional = FALSE)
}
pdf(paste0(pdf_dir, "/IvsM.diffsigPeaks.centromere.VennDiagram.pdf"))
create_proportional_venn(sets=list(IvsM.sigResWald_GR$InterphaseUp$peak_id, 
                                   IvsM.sigResWald_GR$MitosisUp$peak_id,
                                   IvsM.masterpeaks[IvsM.masterpeaks %over% CENPA_OwenSmith_regions]$peak_id, 
                        Interphase_peaks$peak_id, 
                        Mitosis_peaks$peak_id ),
                         set_labels = c("Interphase Up", "Mitotic Up", "Centromeres", "all Interphase peaks", 
                                        "all Mitosis peaks"), 
                         colors = c("#90CAF8FF", "#E7210A", "gold", "#19227EFF", "#CB181D" )
)

create_proportional_venn(sets=list(IvsM.sigResWald_GR$InterphaseUp$peak_id, 
                                   IvsM.sigResWald_GR$MitosisUp$peak_id,
                                   IvsM.masterpeaks[IvsM.masterpeaks %over% CENPA_OwenSmith_regions]$peak_id, 
                        IvsM.masterpeaks$peak_id),
                         set_labels = c("Interphase Up", "Mitotic Up", "Centromeres",
                                        "All accessible peaks"), 
                         colors = c("#90CAF8FF", "#E7210A", "gold", "grey" )
)
dev.off()
```

# extract annotation info of significant peaks
```{r}
anno.IvsM.sigResWald_GR <- lapply(anno.IvsM.sigResWald_GR, as.GRanges)

for(i in 1:length(anno.IvsM.sigResWald_GR)){
anno.IvsM.sigResWald_GR[[i]]$gene_biotype <- gene_annoXeno[match(anno.IvsM.sigResWald_GR[[i]]$geneId, gene_annoXeno$gene_id), ]$gene_biotype
anno.IvsM.sigResWald_GR[[i]]$gbkey <- gene_annoXeno[match(anno.IvsM.sigResWald_GR[[i]]$geneId, gene_annoXeno$gene_id), ]$gbkey
anno.IvsM.sigResWald_GR[[i]]$annotation <- gsub("^\\s*(\\S+)\\s(?!UTR).*", "\\1", anno.IvsM.sigResWald_GR[[i]]$annotation, perl = TRUE)
anno.IvsM.sigResWald_GR[[i]]$Centromere <- ifelse(anno.IvsM.sigResWald_GR[[i]] %over% CENPA_OwenSmith_regions, "Centromeric", "Non-centromeric")
}

# convert anno GR of sig peaks to df 
anno.IvsM.sigResWald_df <- lapply(anno.IvsM.sigResWald_GR, as.data.frame)
anno.IvsM.sigResWald_df$InterphaseUp$SigPeak.Cat <- "Interphase Up"
anno.IvsM.sigResWald_df$MitosisUp$SigPeak.Cat <- "Mitosis Up"
anno.IvsM.sigResWald_df <- rbind(anno.IvsM.sigResWald_df$InterphaseUp, anno.IvsM.sigResWald_df$MitosisUp)
```

# plot peak anno stats
```{r}
# Shapiro-Wilk test for normality
anno.IvsM.masterpeaks_GR <- as.GRanges(anno.IvsM.masterpeaks_GR) # convert masterpeaks anno to GR
shapiro_test <- shapiro.test(sample(abs(anno.IvsM.masterpeaks_GR$distanceToTSS), size=5000))
print(shapiro_test)

# Visualize distribution of differences
ggplot(as.data.frame(anno.IvsM.masterpeaks_GR), aes(x = abs(distanceToTSS))) +
  geom_histogram(binwidth = 10000, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of absolute distance to TSS values", x = "DistancetoTSS", y = "Count")
```
```{r}
# Wilcoxon rank-sum test
w <- wilcox.test(abs(anno.IvsM.sigResWald_GR$InterphaseUp$distanceToTSS), abs(anno.IvsM.sigResWald_GR$MitosisUp$distanceToTSS))

# Kolmogorov–Smirnov test
ks <- ks.test(abs(anno.IvsM.sigResWald_GR$InterphaseUp$distanceToTSS), abs(anno.IvsM.sigResWald_GR$MitosisUp$distanceToTSS))

# plot distribution of peak distance to nearest TSS: no difference I vs M. 
pdf(paste0(pdf_dir, "/distancetoNearestTSS.violinPlot.IvsM.pdf"))
ggplot(anno.IvsM.sigResWald_df, aes(x=SigPeak.Cat, y=abs(distanceToTSS), color=SigPeak.Cat))+geom_violin()+geom_boxplot(width=0.2)+
scale_color_manual(values=c("#E7210A", "#24325FFF"))+
theme_minimal()+
annotate("text", x = 1.5, y = 8e+05, label = paste0("Wilcoxon rank sum p=", w$p.value), vjust = -0.5)+
annotate("text", x = 1.5, y = 8.5e+05, label = paste0("Kolmogorov–Smirnov p=", ks$p.value), vjust = -0.5)  

# Wilcoxon rank-sum test
w <- wilcox.test(abs(anno.IvsM.sigResWald_GR$InterphaseUp$distanceToTSS)[which(abs(anno.IvsM.sigResWald_GR$InterphaseUp$distanceToTSS) <= 50000)], abs(anno.IvsM.sigResWald_GR$MitosisUp$distanceToTSS)[which(abs(anno.IvsM.sigResWald_GR$MitosisUp$distanceToTSS) <= 50000)])

# Kolmogorov–Smirnov test
ks <- ks.test(abs(anno.IvsM.sigResWald_GR$InterphaseUp$distanceToTSS)[which(abs(anno.IvsM.sigResWald_GR$InterphaseUp$distanceToTSS) <= 50000)], abs(anno.IvsM.sigResWald_GR$MitosisUp$distanceToTSS)[which(abs(anno.IvsM.sigResWald_GR$MitosisUp$distanceToTSS) <= 50000)])

ggplot(anno.IvsM.sigResWald_df[abs(anno.IvsM.sigResWald_df$distanceToTSS) <= 50000, ], aes(x=SigPeak.Cat, y=abs(distanceToTSS), color=SigPeak.Cat))+geom_violin()+geom_boxplot(width=0.2)+
scale_color_manual(values=c("#E7210A", "#24325FFF"))+
  scale_y_continuous(limits = c(0,60000))+
  ggtitle("sig peaks found <= 50kb from gene")+
theme_minimal()+
annotate("text", x = 1.5, y =5e+04, label = paste0("Wilcoxon rank sum p=", w$p.value), vjust = -0.5)+
annotate("text", x = 1.5, y = 5.2e+04, label = paste0("Kolmogorov–Smirnov p=", ks$p.value), vjust = -0.5)  
dev.off()
```

# plot type of nearest gene for sig peaks (not much difference I vs M)
```{r}
IvsM.sig.GeneAnno_df <- anno.IvsM.sigResWald_df %>%
  group_by(SigPeak.Cat, gene_biotype, annotation, Centromere) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(SigPeak.Cat, Centromere) %>%
  mutate(PercentTotalPeaks = (n / sum(n))*100)

pdf(paste0(pdf_dir, "/NearestGeneBiotype.IvsM.pdf"))
ggplot(IvsM.sig.GeneAnno_df, aes(x=gene_biotype, y=PercentTotalPeaks, fill=annotation))+
  geom_col()+
  facet_wrap(Centromere~SigPeak.Cat)+
theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Type of nearest gene")+
  scale_fill_cosmic()
dev.off()
```

# extract normalized counts under significant peaks
```{r}
IvsM.NormCountsMat <- counts(IvsM.dds, normalized=T)
sig.IvsM.NormCounts <- IvsM.NormCountsMat[row.names(IvsM.NormCountsMat) %in% sigPeak.ID.Wald, ]
# convert matrix of normalized counts of sig peaks into a dataframe with metaData information for ggplot to accept
sig.IvsM.NormCounts_df <- list()
for(i in 1:nrow(sig.IvsM.NormCounts)){
sig.IvsM.NormCounts_df[[i]] <- data.frame(Counts=sig.IvsM.NormCounts[i, ], sample=names((sig.IvsM.NormCounts[i, ])), Peak_ID= row.names(sig.IvsM.NormCounts)[i], IvsM.metaData)
}
sig.IvsM.NormCounts_df <- do.call(rbind, sig.IvsM.NormCounts_df) # combine list into 1 giant dataframe
```

# convert rlogCounts matrix of significant peaks to df. 
```{r}
sig.IvsM.rlogCounts_df <- list()
for(i in 1:nrow(rlog_IvsM.sigMat)){
sig.IvsM.rlogCounts_df[[i]] <- data.frame(Counts=rlog_IvsM.sigMat[i, ], sample=names((rlog_IvsM.sigMat[i, ])), Peak_ID= row.names(rlog_IvsM.sigMat)[i], IvsM.metaData)
}
sig.IvsM.rlogCounts_df <- do.call(rbind, sig.IvsM.rlogCounts_df) # combine list into 1 giant dataframe
```

# prepare annotation of different genomic regions as dataframe for plotting genome coverage
```{r}
# generate df of centromeric regions
CENPA_df <- as.data.frame(CENPA_OwenSmith_regions)[ ,c(1:3)]
colnames(CENPA_df) <- colnames(hg_gap)[1:3]

# generate df of epigenetic marks
H3K9me3_df <- as.data.frame(H3K9me3_hc_peaks)[ ,c(1:3)]
colnames(H3K9me3_df) <- colnames(hg_gap)[1:3]

H3K27me3_df <- as.data.frame(H3K27me3_hc_peaks)[ ,c(1:3)]
colnames(H3K27me3_df) <- colnames(hg_gap)[1:3]

H3K4me3_df <- as.data.frame(H3K4me3_hc_peaks)[ ,c(1:3)]
colnames(H3K4me3_df) <- colnames(hg_gap)[1:3]

# generate df of chromosome lengths
xenLae2 <- read.table("/lustre/fs4/risc_lab/scratch/jyeung/ncbi-genomes-2022-09-11/Xenopus_laevis_v10.1/GCF_017654675.1_Xenopus_laevis_v10.1_genomic.fna.fai", sep="\t")

# make GRanges object for whole genome
chrom_lengths <- data.frame(Chrom=xenLae2$V1, ChrLength=xenLae2$V2)
chrom_lengths <- chrom_lengths[!grepl("Sca|MT", chrom_lengths$Chrom), ]
```
# prepare dataframe to plot normalized counts under significant peaks
```{r}
# calculate median normalized count under sig peaks within each condition
IvsM.sigResWald_GR <- lapply(IvsM.sigResWald_GR, function(x){
  x$Interphase_normCounts <- apply(sig.IvsM.NormCounts[match(x$peak_id, row.names(sig.IvsM.NormCounts)), 1:3], 1, median)
  x$Mitosis_normCounts <- apply(sig.IvsM.NormCounts[match(x$peak_id, row.names(sig.IvsM.NormCounts)), 4:6], 1, median)
    return(x)}) 
# convert GR to df
sig.IvsM.ResWald_GR_df <-  lapply(IvsM.sigResWald_GR, function(x){
  x <- as.data.frame(x)[ ,c(1:3, 10, 11:12)]
  colnames(x) <- c(colnames(hg_gap)[1:3], "ID", "Interphase_normCounts", "Mitosis_normCounts")
  return(x)})  
# add column for midpoint of peak to plot normalized read count at on the x axis
sig.IvsM.ResWald_GR_df <- lapply(sig.IvsM.ResWald_GR_df, function(x){x %>% 
  mutate(Midpoint = (Start + End) / 2)})
# convert to long format for ggplot to accept
sig.IvsM.ResWald_long_df <- lapply(sig.IvsM.ResWald_GR_df, function(x){x %>%
  dplyr::select(Chrom, Midpoint, Interphase_normCounts, Mitosis_normCounts) %>%
  tidyr::pivot_longer(cols = c(Interphase_normCounts, Mitosis_normCounts), 
                      names_to = "Sample", values_to = "NormCount") %>%
  left_join(chrom_lengths, by = "Chrom")})
```

# plot normalized reads under significant peaks (free y scale for each chromosome)
```{r}
pdf(paste0(pdf_dir, "/GenomeCov.IvsM.sigPeaks.pdf"), width=50, height=100)
# all sig peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$All,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = -20000, ymax = -5000),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$All,
    aes(xmin = Start, xmax = End, ymin = -40000, ymax = -25000),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -60000, ymax = -45000),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -80000, ymax = -65000),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -100000, ymax = -85000),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("All sig peaks")

# interphase up peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$InterphaseUp,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = -3000, ymax = -1000),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$InterphaseUp,
    aes(xmin = Start, xmax = End, ymin = -6000, ymax = -4000),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -9000, ymax = -7000),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -12000, ymax = -10000),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -15000, ymax = -13000),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Interphase Up peaks")

# mitosis up peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$MitosisUp,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = -20000, ymax = -5000),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$MitosisUp,
    aes(xmin = Start, xmax = End, ymin = -40000, ymax = -25000),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -60000, ymax = -45000),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -80000, ymax = -65000),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -100000, ymax = -85000),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Mitosis Up peaks")


# SAME BUT WITHOUT PEAK TRACKS
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$All,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("All sig peaks")

# interphase up peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$InterphaseUp,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Interphase Up peaks")

# mitosis up peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$MitosisUp,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Mitosis Up peaks")

# ONLY PEAK TRACKS: NO GENOME COV
ggplot() +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20, xmax = End+20, ymin = -20, ymax = -5),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$All,
    aes(xmin = Start, xmax = End, ymin = -40, ymax = -25),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -60, ymax = -45),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -80, ymax = -65),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -100, ymax = -85),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("All sig peaks")

# interphase up peaks
ggplot() +
 geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20, xmax = End+20, ymin = -20, ymax = -5),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$InterphaseUp,
    aes(xmin = Start, xmax = End, ymin = -40, ymax = -25),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -60, ymax = -45),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -80, ymax = -65),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -100, ymax = -85),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Interphase Up peaks")

# mitosis up peaks
ggplot() +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = -20, ymax = -5),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$MitosisUp,
    aes(xmin = Start, xmax = End, ymin = -40, ymax = -25),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -60, ymax = -45),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -80, ymax = -65),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -100, ymax = -85),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Mitosis Up peaks")
dev.off()
```

# plot normalized reads under significant peaks (fixed y scale for each chromosome)
```{r}
pdf(paste0(pdf_dir, "/GenomeCov.IvsM.sigPeaks_fixedScale.pdf"), width=30, height=50)
# all sig peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$All,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = -200000, ymax = -50000),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$All,
    aes(xmin = Start, xmax = End, ymin = -400000, ymax = -250000),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -600000, ymax = -450000),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -800000, ymax = -650000),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -100000, ymax = -85000),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free_x", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("All sig peaks")

# interphase up peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$InterphaseUp,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = -3000, ymax = -1000),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$InterphaseUp,
    aes(xmin = Start, xmax = End, ymin = -6000, ymax = -4000),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -9000, ymax = -7000),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -12000, ymax = -10000),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -15000, ymax = -13000),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free_x", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Interphase Up peaks")

# mitosis up peaks
ggplot() +
  geom_point(
    data = sig.IvsM.ResWald_long_df$MitosisUp,
    aes(x = Midpoint, y = NormCount, color = Sample),
    alpha = 0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-200000, xmax = End+200000, ymin = -200000, ymax = -50000),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = sig.IvsM.ResWald_GR_df$MitosisUp,
    aes(xmin = Start, xmax = End, ymin = -400000, ymax = -250000),
    fill = "black", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K4me3_df,
    aes(xmin = Start, xmax = End, ymin = -600000, ymax = -450000),
    fill = "darkorange", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K27me3_df,
    aes(xmin = Start, xmax = End, ymin = -800000, ymax = -650000),
    fill = "olivedrab", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_rect(
    data = H3K9me3_df,
    aes(xmin = Start, xmax = End, ymin = -1000000, ymax = -850000),
    fill = "plum4", alpha = 0.7, inherit.aes = FALSE
  ) +
  # invisible layer to force x-axis length per chromosome
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~ Chrom, scales = "free_x", ncol = 1) +
  theme_classic() +
  xlab("Genomic Position (bp)") +
  ylab("Normalized Counts") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40),
    axis.text.x = element_text(size = 40)
  ) +
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  ggtitle("Mitosis Up peaks")
dev.off()

```

# plot bigwig tracks of epigenetic mods relative to sig peaks & centromeres
```{r}
# generate df of bigwig signal under high confidence peaks
H3K9me3_bw$HistMod <- "H3K9me3"
H3K4me3_bw$HistMod <- "H3K4me3"
H3K27me3_bw$HistMod <- "H3K27me3"

HistMods_bw_df <- as.data.frame(c(H3K9me3_bw, H3K4me3_bw, H3K27me3_bw))[ ,c(1:3, 6:7)]
colnames(HistMods_bw_df) <- c(colnames(hg_gap)[1:3], "score", "HistMod")
HistMods_bw_df$Midpoint <- (HistMods_bw_df$Start+HistMods_bw_df$End)/2


# bin 20bp intervals into 2kb intervals instead
HistMods_bw_2kb <- HistMods_bw_df %>%
  mutate(Bin2kb = floor(Start / 2000) * 2000) %>%   # bin start into 1kb blocks
  group_by(Chrom, HistMod, Bin2kb) %>%
  summarise(
    score = mean(score, na.rm = TRUE),
    Start = min(Start),
    End   = max(End),
    Midpoint = (min(Start) + max(End)) / 2,
    .groups = "drop"
  )
```

# plot histone mods signal vs norm counts under sig peaks
```{r}
pdf(paste0(pdf_dir, "/GenomeCov.IvsM.sigPeaks.vs.HistModSignal.pdf"), width=50, height=100)
ggplot() +
  geom_col(
    data = HistMods_bw_2kb,
    aes(x = Midpoint, y = -score, fill = HistMod), position="dodge", alpha=0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = 0, ymax = 8),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_point(
    data = sig.IvsM.ResWald_long_df$All,
    aes(x = Midpoint, y=(NormCount/1e+04)+10, color=Sample), alpha=0.5, size=2
  ) +
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~Chrom, scales = "free_x", ncol=1) +
  theme_classic() +
  xlab("Genomic Position") +
  ylab("Bigwig score : Normalized Counts/10^4") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40), 
    axis.text.x = element_text(size = 40), 
    axis.title = element_text(size = 40),
    title=element_text(size = 40), 
    legend.text=element_text(size = 40)
  )+
  scale_fill_manual(
    values = c("olivedrab", "darkorange", "plum4")
  )+
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  scale_y_continuous(limits = c(-50, 50), oob = oob_squish)+ # for values above scale limits, just squish to max limit and do not remove
  ggtitle("norm counts under all sig peaks vs histone mods signal")

# interphase up peaks 
ggplot() +
  geom_col(
    data = HistMods_bw_2kb,
    aes(x = Midpoint, y = -score, fill = HistMod), position="dodge", alpha=0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = 0, ymax = 8),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_point(
    data = sig.IvsM.ResWald_long_df$InterphaseUp,
    aes(x = Midpoint, y=(NormCount/1e+02)+10, color=Sample), alpha=0.5, size=2
  ) +
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~Chrom, scales = "free_x", ncol=1) +
  theme_classic() +
  xlab("Genomic Position") +
  ylab("Bigwig score : Normalized Counts/10^2") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40), 
    axis.text.x = element_text(size = 40), 
    axis.title = element_text(size = 40),
    title=element_text(size = 40), 
    legend.text=element_text(size = 40)
  )+
  scale_fill_manual(
    values = c("olivedrab", "darkorange", "plum4")
  )+
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  scale_y_continuous(limits = c(-50, 50), oob = oob_squish)+ # for values above scale limits, just squish to max limit and do not remove
  ggtitle("norm counts under Interphase Up peaks vs histone mods signal")

# mitosis up peaks 
ggplot() +
  geom_col(
    data = HistMods_bw_2kb,
    aes(x = Midpoint, y = -score, fill = HistMod), position="dodge", alpha=0.5
  ) +
  geom_rect(
    data = CENPA_df,
    aes(xmin = Start-20000, xmax = End+20000, ymin = 0, ymax = 8),
    fill = "gold", alpha = 0.7, inherit.aes = FALSE
  ) +
  geom_point(
    data = sig.IvsM.ResWald_long_df$MitosisUp,
    aes(x = Midpoint, y=(NormCount/1e+04)+10, color=Sample), alpha=0.5, size=2
  ) +
  geom_blank(
    data = chrom_lengths,
    aes(x = ChrLength, y = 0)  # ensures each facet spans full chromosome
  ) +
  facet_wrap(~Chrom, scales = "free_x", ncol=1) +
  theme_classic() +
  xlab("Genomic Position") +
  ylab("Bigwig score : Normalized Counts/10^4") +
  theme(
    strip.text = element_text(size = 40),
    axis.text.y = element_text(size = 40), 
    axis.text.x = element_text(size = 40), 
    axis.title = element_text(size = 40),
    title=element_text(size = 40), 
    legend.text=element_text(size = 40)
  )+
  scale_fill_manual(
    values = c("olivedrab", "darkorange", "plum4")
  )+
  scale_color_manual(values = c("#E7210A", "#24325FFF"))+
  scale_y_continuous(limits = c(-50, 50), oob = oob_squish)+ # for values above scale limits, just squish to max limit and do not remove
  ggtitle("norm counts under mitosis up peaks vs histone mods signal")
dev.off()
```
